Guidelines and Policy for the Dark Energy Survey Strong Lensing Science Team using \SW

### 2015-03-30 doc/sw-publication-policy-des.tex

\begin{abstract} 
\noindent \SW was conceived as a strong lens
discovery {\it service} by harnessing the power of citizen scientists that the DES strong lensing group (hereafter DES survey science team or DES SST) can make use of
(within the logistical constraints). In this document we discuss how the DES SST can become involved with \SW and the relationship
between the DES survey science team and the \SW collaboration. 
This document includes guidelines for proposing projects, data provision, data policy and for publications of \SW-enabled discoveries.
\end{abstract}

@@

\section{Introduction to \SW}

@@

\section{Defining a suitable \SW science project}
\label{sec:project}

@@

\textit{Targeted} searches are also possible, as we performed in the second \SW project on the VICS82 survey area. Here we inspected $\sim$40,000 images from the VICS82 survey centred on likely strong lens targets, e.g. LRGs, groups, clusters, quasars.  Again this was more images than the SST could manage to inspect alone.

@@

\subsection{DES specific thoughts}
In the case of DES, with its long data release plan and large strong lens group, we request the SST to consider the following:
\begin{enumerate}
\item The \SW facility should be used for science projects that cannot be completed by the Strong Lens Group itself. 
\item Any \SW search should be complementary, not in competition, to the lens finding being conducted by the DES SST itself.
\item While \SW uses citizen science, its primary goal is \textbf{not outreach}, it is harnessing citizen power for visual classifications (and eventually modelling) powered by citizens to perform science.
\item A duplicate project is not encouraged provided there is sufficient scientific motivation and justification to do so.
\item \SW requires volunteers/citizens to give up their personal time to conduct the experiment, this should not be taken lightly, the crowd are motivated by contributing to real science not just doing things for the sake of them. This is the reason the suitability of the project for citizen scientists should be carefully considered. The \SW PIs will be able to discuss any planned \SW-DES projects with the DES SST.
\item The success of \SW-DES will rely on input from the DES SST working with the \SW collaboration, dedicated effort is required for this and the DES SST should identify team members who can commit time to the \SW-DES tasks for this before embarking on a project. The \SW PIs will help to establish the main efforts required, these will depend on the actual project put forward by the DES SST (e.g. image preparation, simulations).
\end{enumerate}

@@

Please do remember that \SW is setup to provide a service to help your team find lenses, any discovery in \SW-DES is the DES SSTs discovery. The goal of the \SW PIs or the \SW collaboration is to enable your science project and make it work. Using SWAP we will send the DES SST lists of the candidates selected by the citizens (at some regular intervals or at the end of a project). If the DES SST wish to have the \SW Pis input or to collaboratively work on targets the \SW team are happy to become more involved (the \SW Pis have strong lensing science interests) but we will not do anything with the sources found through \SW other than provide it to the DES SST.  We are happy to then work with you as you wish.

The primary duty of SW PIs is to enable the discovery of lenses through citizen science and forward the catalog of candidates to the SST (the latter part essentially happens automatically through the pipeline SWAPR). The SW team is happy to help with further grading of the candidates or assist with any scientific investigations and follow-up of the candidates, if the DES SST so requests.

Note all possible candidates are in the public domain as all images are available online (but without WCS information). These targets may be discussed by citizens on \SW\,\Talk. Members of the DES SST are expected to engage in discussions with citizens through \Talk, which in turn has the very positive effect of motivating the citizens to contribute further.

\subsection{Previewing images}
We are aware that there is some interest in the DES SST for using \SW web interface to preview images for the SST themselves. As this is not the function of the \SW interface or the Zooniverse, this would only be possible under certain conditions:
\begin{enumerate}[label=(\alph*)]
\item the images\footnote{can still be stripped of WCS information (please see Section \ref{sec:data}} inspected are destined to be seen by the crowd  \textbf{in parallel} with the SST, therefore the URL for the inspection will be publicly available.;
\item the framework of the image presentation (i.e the \SW interface) is the same as the interface presented to the volunteers (image clicks only and no comments or ranking possible);
\item all SST inspectors would have to register with the Zooniverse to classify.
\end{enumerate}

@@

In our experience, a citizen science project like SW cannot function successfully without significant support and co-ordination of the SST. Thus, we recommend the all DES members with an nterest in strong lensing are told about \SW and how it can complement their lens finding effort. We are here to assist your science! While we can't expect the entirety of the DES Strong Lensing Group or the STRIDES consortium to be behind \SW-DES, all should be made aware of the projects being undertaken and planned with \SW.

In particular, for the first potential \SW-DES project (probably on the Y1 data) we encourage the DES SST to strongly consider the timing for when they wish to ingest the images into \SW. This could be delayed until the SST has performed all the visual inspection or automated algorithms they wish to to avoid any issue of competition with \SW (if that is a concern).

@@

\begin{itemize}
\item Survey area
\item Bands to be included
\item Total number of images to be viewed (approx.)
\item Targeted survey or blind field search \footnote{Targeted: inspection of a list of candidates; Blind: inspection of tiles of a survey area}
\item Brief outline of the overall goal(s) of that inspection, highlighting how this is complementary (not competing) with the goals of the DES lens finding group
\item Proprietary data issues 
\item identify a core team who will lead and work closely with \SW
\end{itemize}

@@

\section{Data Preparation \& \SW interface}
\label{sec:data}

@@

\item In {\bf running the \SW analysis pipeline} which will be our responsibility. This will provide candidates to be sent o the DES SST.

@@

\subsection{Journal Papers}
\label{sec:papers}
In return for the work outlined in Section \ref{sec:data} in setting up the new surveys, the \SW PIs will nominate a
small number of core \SW collaboration members as authors on \textbf{papers that
present discovery of a lens or sample of lenses enabled by the \SW system}. A ``discovery'' is defined by the announcement of the source(s) to the astronomical community in a journal paper. We would expect the \SW author list to be a small number (less than ten), primarily the Principal Investigators, a few others working on the ingestion of the new data or moderating \Talk. All post-discovery follow-up papers would of course be exempt from this: the idea is that \SW is purely a gravitational lens discovery service, but not one that runs by magic.

We request that authors of \SW-enabled discovery papers should circulate drafts to the list of \SW authors and allow a minimum of two weeks to give time for comments. We also request that any papers based on \SW enabled discoveries should cite the \SW system papers. This policy document will be updated with the correct citations to the \SW system papers when available.

>>>>

\begin{abstract} 
\noindent \SW was conceived as a strong lens
discovery {\it service} by harnessing the power of citizen scientists that the DES strong lensing group (hereafter DES survey science team or DES SST) can make use of
(within the logistical constraints). In this document we discuss how the DES SST can become involved with \SW and the relationship
between the DES SST and the \SW collaboration. 
This document includes guidelines for proposing projects, data provision, data policy and for publications of \SW-enabled discoveries.
\end{abstract}

@@

\section{Introduction to \sw}

@@

\section{Defining a suitable \sw science project}
\label{sec:project}

@@

\textit{Targeted} searches are also possible, as we performed in the second \SW project on the VICS82 survey area. Here we inspected $\sim$40,000 images from the VICS82 survey centred on likely strong lens targets, e.g. LRGs, groups, clusters, quasars.  Again this was more images than the VICS82-SST could manage to inspect alone.

@@

\subsection{DES specific thoughts}
In the case of DES, with its long data release plan and large strong lens group, we request the SST to consider the following:
\begin{enumerate}
\item The \SW facility should be used for science projects that cannot be completed by the Strong Lens Group itself. 
\item Any \SW search should be complementary, not in competition, to the lens finding being conducted by the DES SST itself.
\item While \SW uses citizen science, its primary goal is \textbf{not outreach}, it is harnessing citizen power for visual classifications (and eventually modelling) to perform science.
\item A duplicate project is not encouraged provided there is sufficient scientific motivation and justification to do so.
\item \SW requires volunteers/citizens to give up their personal time to conduct the experiment, this should not be taken lightly, the crowd are motivated by contributing to real science not just doing things for the sake of them. This is the reason the suitability of the project for citizen scientists should be carefully considered. The \SW PIs will be able to discuss any planned \SW-DES projects with the DES SST. It is important to bear this in mind for long-term surveys like DES, if needed the crowd needs to be engaged and feel like they are contributing while the entire \SW-DES project runs.
\item The success of \SW-DES will rely on input from the DES SST working with the \SW collaboration, dedicated effort is required for this and the DES SST should identify team members who can commit time to the \SW-DES tasks for this before embarking on a project. The \SW PIs will help to establish the main efforts required, these will depend on the actual project put forward by the DES SST (e.g. image preparation, simulations, interacting with volunteers on \Talk).
\end{enumerate}

@@

Please do remember that \SW is setup to provide a service to help your team find lenses. Any discovery in \SW-DES is the DES SSTs discovery. The goal of the \SW PIs or the \SW collaboration is to enable your science project and make it work. Using the automatic analysis pipeline (SWAPR) we will send the DES SST lists of the candidates selected by the citizens (at some regular intervals or at the end of a project).

If the DES SST wish to have the \SW PIs input or to collaboratively work on targets the \SW team are happy to become more involved (the \SW PIs have strong lensing science interests).  
The \SW team is happy to help with further grading of the candidates or assist with any scientific investigations and follow-up of the candidates, if the DES SST so requests.

\subsection{Previewing images and \Talk}
We are aware that there is some interest in the DES SST for using the \SW web interface to preview images for the SST themselves. This is not the function or purpose of the \SW interface or the Zooniverse. However  it will be possible for DES SST members to view the images in parallel with the crowd through the public interface.  The DES SST members who wish to do this will need to register with the Zooniverse to classify. We can extract the classifications of SSTs only to provide and expert selected sample, if required. The images marked by any inspector (SST or citizen)  are automatically recorded in an individual's \SW/Zooniverse account. These can be accessed through the \SW interface and as a 'collection' in \Talk. This collection can be explored (and discussed if required) through the \Talk interface.

\SW \Talk is a key and integral part of \SW. The DES SST are strongly encouraged to interact with the volunteers on \Talk. This exchange allows the crowd to learn from the experts, this in turn has the very positive effect of motivating the citizens to contriute further. In addition it also results in better and informed classification when they return to that task.

Note that all possible candidates are in the public domain as all images are available online (but without WCS information).

@@

In our experience, a citizen science project like SW cannot function successfully without significant support and co-ordination of the SST. Thus, we recommend that all DES members with an interest in strong lensing are told about \SW and how it can complement their lens finding effort. We are here to assist your science! While we can't expect the entirety of the DES Strong Lensing Group or the STRIDES consortium to be behind \SW-DES, all should be made aware of the projects being undertaken and planned with \SW.

We encourage the DES SST to strongly consider the timing for when they wish to ingest the images into \SW so that it is in line with the interests of the DES SST and consortium.

@@

\begin{itemize}
\item Survey area
\item Bands to be included
\item Total number of images to be viewed (approx.)
\item Targeted survey or blind field search \footnote{Targeted: inspection of a list of candidates; Blind: inspection of tiles of a survey area}
\item Brief outline of the overall goal(s) of that inspection, highlighting how this is complementary (not competing) with the goals of the DES lens finding group
\item Proprietary data issues 
\item identify an initial core team who will lead and work closely with \SW
\end{itemize}

@@

\section{Data Preparation \& the \SW interface}
\label{sec:data}

@@

\item In {\bf running the \SW analysis pipeline} which will be our responsibility. This will provide citizen selected candidates to be sent to the DES SST.

@@

\subsection{Journal Papers}
\label{sec:papers}
In return for the work outlined in Section \ref{sec:data} in setting up the new surveys, the \SW PIs will nominate a core \SW collaboration members as authors on \textbf{papers that
present discovery of a lens or sample of lenses enabled by the \SW system}. A ``discovery'' is defined by the announcement of the source(s) to the astronomical community in a journal paper. We would expect the \SW author list to be a small number (approx. ten), primarily the Principal Investigators, a few others working on the ingestion of the new data or moderating \Talk. All post-discovery follow-up papers would of course be exempt from this (unless individuals from \sw team had contributed sufficiently such that the lead author wished to include them).

We request that authors of \SW-enabled discovery papers should circulate drafts to the list of \SW authors and, if possible, allow a minimum of two weeks to give time for comments. We also request that any papers based on \SW enabled discoveries should cite the \SW system papers. This policy document will be updated with the correct citations to the \SW system papers when available.
